---
title: "SFG BigData Workshop"
subtitle: "RStudio conference 2018 de-brief"
author: "JC"
date: "March, 1, 2018"
output:
  xaringan::moon_reader:
      nature:
        highlightStyle: github
        highlightLines: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 3)
```

# Outline

- Basics of a connection

- BigQuery stuff

- `sparklyr`
  - visualization
  - model fitting
  - model testing

---

# Packages I am using

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(DBI)
  library(bigrquery)
  library(dbplot)
  library(sf)
  library(tmap)
  library(ggplot2)
  library(sparklyr)
})
```


---

# Establishing a connection

```{r}
BQc <- bigrquery::dbConnect(drv = bigrquery::dbi_driver(), 
                      project = "ucsb-gfw", 
                      dataset = "sfg_bigdata", 
                      allowLargeResults = TRUE)
```

---

# What's in there?

```{r}
DBI::dbListTables(BQc)
```

---

# Get first 5 records of vessel name, and position

with `odbc::dbGetQuery`

```{r}
odbc::dbGetQuery(BQc,"SELECT
  vessel_name, lon, lat
FROM
  [ucsb-gfw:sfg_bigdata.chile_gfw]
LIMIT
  5")
```

---

# Get first 5 records of vessel name, and position

in `SQL` chunk

```{sql, connection = BQc, output.var = "pos"}
SELECT
  vessel_name, lon, lat
FROM
  [ucsb-gfw:sfg_bigdata.chile_gfw]
LIMIT
  5
```

```{r}
pos
```

---

# Let's try with `dplyr`

```{r}
chile_gfw <- dplyr::tbl(BQc, "chile_gfw")
```

`chile_gfw` is now on our local environment... but not the whole data... so what is it?

```{r}
class(chile_gfw)
```

```{r, message = T, warning = T}
dplyr::show_query(chile_gfw)
```

```{r}
dim(chile_gfw)
```

---

Same from objects I get out of it:

```{r}
don_julio <- chile_gfw %>% 
  filter(vessel_name == "DON JULIO")
```

```{r, message = T, warning = T}
show_query(don_julio)
```


---

# Get first 5 records of vessel name, and position

with `dplyr`

```{r, message = T, warning = T}
chile_gfw %>% 
  select(vessel_name, lon, lat) %>% 
  head(5) %>% 
  show_query()
```

---

# Get first 5 records of vessel name, and position

with `dplyr`

```{r}
chile_gfw %>% 
  select(vessel_name, lon, lat) %>% 
  head(5)
```

---

# `!!`
## Bang-bang to translate `r`-specific code

```{r, message = T, warning = T}
chile_gfw %>%
  select(vessel_name) %>% 
  mutate(today = Sys.Date()) %>% #<<
  show_query()
```

---
# `!!`
## Bang-bang to translate `r`-specific code

```{r, message = T, warning = T}
chile_gfw %>%
  select(vessel_name) %>% 
  mutate(today = !!Sys.Date()) %>% #<<
  show_query()
```

**Do not abuse the `!!`**

---
class: inverse, center, middle

# `dplyr` is way easier!

## ... or at least more familiar to us

---

# Let's run a usual question

Who fishes more?

```{r, warning = F, message = F}
chile_gfw %>% 
  filter(nnet_score == 1L) %>% 
  group_by(vessel_name) %>% 
  summarize(total_fishing = sum(hours, na.rm = T)/24L) %>% 
  arrange(desc(total_fishing)) %>% 
  head(5)
```

---

# Adding data from disk

Read it into R
```{r}
trip_spp_landings <- readRDS("data/trip_spp_landings.rds")
```

Send it up
```{r, eval = F}
DBI::dbWriteTable(conn = BQc,
                  name = "trip_spp_landings",
                  value = trip_spp_landings,
                  temporary = TRUE)
```

We need to have a `tbl` in here

```{r}
tbl_spp_landings <- tbl(BQc, "trip_spp_landings")
```

---
class: inverse, center, middle

# I want to have my summarized data on disk

---

# Enter `collect`

- `collect` gives you the data, rather than the sql statement or the `tbl` object

```{r}
don_julio_collected <- don_julio %>% 
  filter(nnet_score == 1) %>% 
  mutate(lon = floor(lon*2)/2 + 0.25,
         lat = floor(lat*2)/2 + 0.25) %>% 
  group_by(lon,lat) %>% 
  summarize(total_fishing = sum(hours, na.rm = T)/24L) %>% 
  collect()
```

```{r}
class(don_julio_collected)
```

```{r}
dim(don_julio_collected)
```


---

# Visualization of collected data

```{r, echo = F}
proj <- "+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs"
data("World")
cone <- World[World$name %in% c("Chile", "Argentina", "Peru", "Bolivia"),] %>% 
  as("sf") %>% 
  st_transform(proj)
```

```{r, fig.height = 3}
ggplot() +
  geom_sf(data = cone) +
  geom_raster(data = don_julio_collected,
              aes(x = lon, y = lat, fill = total_fishing)) +
  scale_fill_gradientn(colours = colorRamps::matlab.like(20)) +
  theme_bw()
```


---
class: inverse, center, middle

# `sparklyr`

---

# Connecting

Establishing a connections is quite straightforward

```{r}
sc <- spark_connect(master = "local")
```

--
 
## Look at your `Connections` pane in RStudio

---

#"Map" data

```{r}
columns <- read.csv("data/flights/flights_2013_1.csv", nrows = 5, stringsAsFactors = F) %>% 
  rename_all(tolower) %>%
  purrr::map(function(x) "character")
```

```{r}
head(columns)
```

---

## Use spark_read(_csv) to “map” the file’s structure and location to the Spark context

```{r}
spark_flights <- spark_read_csv(
  sc,
  name = "spark_flights",
  path = "./data/flights/",
  memory = FALSE,
  columns = columns,
  infer_schema = FALSE
)
```

--

## Look at your `Connections` pane in RStudio

---

# What's in there?

```{r}
class(spark_flights)
```

```{r}
dim(spark_flights)
```

--

Hu, the same stuff as before

---
class: inverse, center, middle

# Common question at the RStudio conference:

## Will it pipe?

---

# `dplyr-able`

```{r}
spark_flights %>% 
  count()
```

# Yes!

---

# Things to have in mind

- Some `tidyr` verbes don't work with `sdf`'s
  - `spread`
  - `gather`
- But:
  - There are [`sdf_` transformers](http://spark.rstudio.com/reference/#section-spark-dataframes)

---

# Cache the data

## Open spark UI in `Connections` pane

--

```{r}
cached_flights <- spark_flights %>% 
  select(month,
         dep_time,
         arr_time,
         arr_delay,
         dep_delay,
         distance,
         sched_dep_time,
         sched_arr_time) %>%
  mutate_all(as.numeric) %>% 
  compute("cached_flights") #<<
```

--

## Look at teh connections pane again

---

# Confirm it works

```{r}
cached_flights %>% 
  count()
```

```{r}
head(cached_flights, 5)
```


---

# Visualizing with `dbplot`

```{r}
cached_flights %>%
  dbplot_line(month) +
  theme_bw()
```

---

# Fit a model

## Prepare the data

```{r}
sample_data <- cached_flights %>%
  filter(!is.na(arr_delay)) %>%
  ft_binarizer(input.col = "arr_delay", #<<
               output.col = "delayed", #<<
               threshold = 15) %>% #<<
  ft_bucketizer(input.col = "sched_dep_time", #<<
                output.col = "dephour", #<<
                splits = c(0, 400, 800, 1200, 1600, 2000, 2400)) %>% #<<
  mutate(dephour = paste0("h", as.integer(dephour))) %>%
  sdf_partition(training = 0.01, testing = 0.09, other = 0.9)
```

---

# Make sure it works

```{r}
tally(sample_data$training)
```

```{r}
training <- sdf_register(sample_data$training, "training")
tbl_cache(sc, "training")
```

---

# Fit a model

Use `sparklyr::ml_logistic_regression`

```{r}
delayed_model <- training %>%
  ml_logistic_regression(delayed ~ dep_delay + dephour)
```

```{r}
summary(delayed_model)
```

```{r}
class(delayed_model)
```

---

# Run predictions in spark

```{r}
delayed_testing <- sdf_predict(delayed_model, sample_data$testing)
```

```{r}
delayed_testing %>% 
  select(delayed, dephour, prediction, probability) %>% 
  head(5)
```

---

# How did we do?

```{r}
delayed_testing %>%
  group_by(delayed, prediction) %>%
  tally()
```

---
class: inverse, center, middle

The end!

---


# Secure connections: 5 ways

- `config`

- keyring

- Environment vaiables

- `options`

- Prompt for credentials

---

# `config.yml` file with structure:


.pull-left[
```
default:
  datawarehouse-dev:
    driver: 'BigQueryDriver'
    server: 'localhost'
    uid: 'rstudio_admin'
    pwd: 'admin_user_be_careful'
    port: 5432
    database: 'postgres'
```
]

.pull-right[
```
con <- DBI::dbConnect(odbc::odbc(),
                      Driver = dw$driver,
                      Server = dw$server,
                      UID    = dw$uid,
                      PWD    = dw$pwd,
                      Port   = dw$port,
                      Database = dw$database)
```
]


